---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



# kneedle

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/etam4260/kneedle/branch/main/graph/badge.svg)](https://codecov.io/gh/etam4260/kneedle?branch=main)
<!-- badges: end -->

The goal of kneedle is to provide an easy to use implementation of the kneedle 
algorithm developed at https://raghavan.usc.edu/papers/kneedle-simplex11.pdf. Rather than
following an object oriented implementation (KneeLocator in Python and Kneer) of the
algorithm, this takes a more functional approach which is easier to understand
for beginners. The main functionality is to detect an inflection point, 
more commonly associated in continuous function, in discrete datasets which closely
follow an exponential decay or growth or logarithmic decay or growth.

Please report bugs to Emmet Tam, at emmet_tam@yahoo.com or through github.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("etam4260/kneedle")
```
## Example

This is a basic example:

```{r example}
library(kneedle)

# The base function 'guesses' whether the data is increasing or decreasing.
# (Will assume if the derivative of the first and last point is positive or negative:
# positive is increasing and negative decreasing)
# It also guesses the 'concavity' of the data. (Will use second derivative calculations
# using all the points and then taking the mean of those values: if the mean is >= 0, then
# it is concave; if mean is < 0, then is convex)
# If no sensitivity value provided, it will default to a sensitivity of 1. 

# First and second inputs must be vectors of integers or doubles.

knee <- kneedle(c(1,2,3,4,5,6), c(0,1,2,3,40,100))
print(knee)
plot(c(1,2,3,4,5,6), c(0,1,2,3,40,100), xlab = "x", ylab = "y", pch=21, col="blue", bg="lightblue", type = "b")
```

## Parameters
This describes a few of the parameters that can be changed.

```{r parameters}
library(kneedle)
# Concavity parameter can be changed if the graph is convex. Furthermore, you 
# specify if the y is increasing or decreasing as x increases. However, the
# algorithm has 'autodetection' for those values.

knee <- kneedle(c(1,2,3,4,5), c(0,20,40,41,42), concave = FALSE, decreasing = FALSE)
print(knee)
plot(c(1,2,3,4,5), c(0,20,40,41,42), xlab = "x", ylab = "y", pch=21, col="blue", bg="lightblue", type = "b")


knee <- kneedle(c(1,2,3,4,5), c(100,99,98,50,0), concave = FALSE, decreasing = TRUE)
print(knee)
plot(c(1,2,3,4,5), c(100,99,98,50,0), xlab = "x", ylab = "y", pch=21, col="blue", bg="lightblue", type = "b")
```


## Sensitivity
This describes how the sensitivity parameter affects the output.

```{r sensitivity}
library(kneedle)
# Sensitivity defaults to 1 as per the referenced paper. However, you can adjust 
# it. A higher sensitivity make the rules more 'stringent' in classifying a 
# 'candidate knee' point as a knee. 

knee <- kneedle(c(1,2,3,4,5), c(0,1,2,40,60), sensitivity = 1)
print(knee)
plot(c(1,2,3,4,5), c(0,1,2,40,60), xlab = "x", ylab = "y", pch=21, col="blue", bg="lightblue", type = "b")

# In this case with sensitivity = 2, we see that that (3,2) is no longer considered a knee point. No
# other knees were detected.

knee <- kneedle(c(1,2,3,4,5), c(0,1,2,40,60), sensitivity = 2)
print(knee)
plot(c(1,2,3,4,5), c(0,1,2,40,60), xlab = "x", ylab = "y", pch=21, col="blue", bg="lightblue", type = "b")
```
